### 2016-11-29 paper/paper.tex
more text on VAE

\paragraph{Variational Autoencoders\cite{vae:2013}}
\label{par:overview_vae}
Autoencoders comprise of two components, the encoder and decoder
where the encoder takes input data and translates it into a latent structure while
the decoder takes the latent structure and produces data which is as close as possible to the initial input data.

In the case of VAEs, the distribution is intractable and we can only approximate the posterior distribution.
% intractable -> use neural network, which we can train
% --> no closed form possible
We use the KL-Divergence to train the network.

% TODO read chapter 19, approximate inference
In VAEs however, this

@@

>>>>

\paragraph{Variational Autoencoders\cite{vae:2013}}
\label{par:overview_vae}
Variational Autoencoders (VAE) use two networks called encoder and decoder where the encoder transform the given data into a latent space and the decoder translates this latent space back into the data space.
The input and output space are usually high-dimensional data spaces, for example images while the latent space is low-dimensional in comparison.
Like other autoencoders, VAEs try to reconstruct the input data using the latent variables as closely as possible.
But in addition to regular autoencoders encourages the VAE the latent space to match a specific probability distribution.
This encouragement is done using a variational lower bound on the distribution output by the encoder network.

%where the encoder takes input data and translates it into a latent structure while
%the decoder takes the latent structure and produces data which is as close as possible to the initial input data.

%In the case of VAEs, the distribution is intractable and we can only approximate the posterior distribution.
% intractable -> use neural network, which we can train
% --> no closed form possible
%We use the KL-Divergence to train the network.

% TODO read chapter 19, approximate inference
%In VAEs however, this

@@

\subsection{Loss function}
The loss function of the VAE consists of two terms, the negative log-likelihood or the reconstruction error $\mathcal{C}_{\mathrm{rec}}$ and the second term $\mathcal{C}_{\mathrm{KL}}$ serves as the regularizer.
We will discuss both terms in detail below.

\paragraph{Reconstruction error}
$\mathcal{C}_{\mathrm{rec}}$ measures how well the decoder reconstructs the input data using the latent space $z$.
This error is measures using the negative log-likelihood of the conditional probability distribution $p_\theta(x_i|z)$ where $z$ is sampled from $q_\varphi(z | x_i)$.
$\mathcal{C}_{\mathrm{rec}}$ is necessary to force the encoder to produce latent variables which can be used to reconstruct the input data well.

\paragraph{KL divergence (regularization)}
The second term in $\mathcal{C}$ is the KL divergence between the approximated distribution $q_\varphi(z|x)$ and $p(z)$. This serves as the regularizer in the VAE as it encourages $q_\varphi(z|x)$ to match $p(z)$.\\
% move to architecture:
%In the original VAE paper, the unit gaussian distribution with diagonal variance was proposed for $p(z)$ and although there have been multiple proposals (citations, multi-modal etc) for different distributions we will stick with $\mathcal{N}(0,I)$ for the sake of simplicity.
% ---

For two probability distributions $P$ and $Q$, the KL divergence $\mathcal{D}_{\mathrm{KL}}(P||Q)$ can informally be described as the amount of information which is lost when using $Q$ to represent $P$.
The KL divergence does not obey the triangle inequality and is also not symmetrical, therefore it does not qualify as a metric.
$\mathcal{D}_{\mathrm{KL}}(P||Q)$ can nevertheless be understood as a measure of the difference between $P$ and $Q$ and is as such used in the VAE to let $q_\varphi(z|x)$ approach $p(z)$.
